---
title: "Process and clean data"
output: html_notebook
---


# Load libraries
```{r}
library(readr)
library(dplyr)
library(rvest) #scrape
library(stringr) #pad
library(clock) #under_minutes
library(ggplot2)
```

# Scrape/download variable name table
```{r}
# download variable name table
u = "https://www.transtats.bts.gov/DL_SelectFields.aspx?gnoyr_VQ=FGK&QO_fu146_anzr=b0-gvzr"

css_selector = "#myTable"
var_info_df = u %>% 
  read_html() %>% 
  html_element(css = css_selector) %>%
  html_table(fill = TRUE) 
colnames(var_info_df) = var_info_df[1,]
var_info_df = var_info_df[-1,-3] # remove column names row and support table column

# remove subheadings from table 
subheading_ind = c(1,3,9,24,34,44,54,64,69,76,82,86)
var_info_df = var_info_df[-subheading_ind,]

# save df 
write.csv(var_info_df, "Data/var_info_df.csv", row.names = F)
```

# Load Flight data 
```{r}
df = read.csv("Data/march2021.csv")
colnames(df) = var_info_df[["Field Name"]]
```

# Preprocessing EDA
```{r}
# dataset structure
str(df)

# identify variables with one response - we can remove these non-meaningful variables
colnames(df[,sapply(df, function(x) length(unique(x))) == 1]) 

# explore missing values NAs and plot percentage missing for each column
missing_df = data.frame("var_name" = colnames(df),
                    "missing_pct" = colSums(is.na(df)) / nrow(df))
ggplot(missing_df, aes(x=reorder(var_name,-missing_pct),y=missing_pct)) +
  geom_bar(stat="identity",fill="blue")+
  coord_flip()+theme_bw()
```

After doing EDA on the pre-processed data, we have identified tasks to complete during during the data cleaning:  

1. correct column types for nominal, time, date, and numeric features.

2. drop the columns with one response. 

3. drop columns with majority missing values

# Because we have so many variables, we can explore them in sections
```{r}
# html table contained subheadings, create subsets for each data category 
time_df = var_info_df[1:6,]
airline_df = var_info_df[7:20,]
origin_df = var_info_df[21:29,]
destination_df = var_info_df[30:38,]
depart_df = var_info_df[39:47,]
arrive_df = var_info_df[48:56,]
canc_div_df = var_info_df[57:60,]
summary_df = var_info_df[61:66,]
delay_cause_df = var_info_df[67:71,]
gate_return_df = var_info_df[72:74,]
diverted_air_df = var_info_df[75:nrow(var_info_df),]
```

# clean each data subset 
```{r}
# create empty lists to store variables which need type change
# we will drop all variables not stored in one of these lists
nominal_vars = c()
time_vars = c()
date_vars = c()
num_vars = c()
```

## Time Period
```{r}
df[, time_df[[1]]] %>% head(2)

nominal_vars = c(nominal_vars, 
                 c("Year", "Quarter", "Month", "DayofMonth", "DayOfWeek"))
date_vars = c(date_vars, "FlightDate")
```

## Airline
```{r}
df[, airline_df[[1]]] %>% head(2)

# to remove exact collinearity, we will not include all airline identification columns
nominal_vars = c(nominal_vars,
                 c("Marketing_Airline_Network", "Operating_Airline", 
                 "Tail_Number", "Flight_Number_Operating_Airline"))
```

## Origin
```{r}
df[, origin_df[[1]]] %>% head(2)

# to remove exact collinearity, we will not include all origin identification columns
nominal_vars = c(nominal_vars, c("Origin", "OriginCityName", "OriginState"))
```

## Destination
```{r}
df[, destination_df[[1]]] %>% head(2)

# to remove exact collinearity, we will not include all destination identification columns
nominal_vars = c(nominal_vars, c("Dest", "DestCityName", "DestState"))
```

## Departure Performance
```{r}
df[, depart_df[[1]]] %>% head(2)

# will not include DepDelayMinutes as equivalent to DepDelay: with negatives instead of zeros 
nominal_vars = c(nominal_vars, c("DepDel15", 
                                 "DepartureDelayGroups", "DepTimeBlk"))
time_vars = c(time_vars, c("CRSDepTime", "DepTime", "WheelsOff"))
num_vars = c(num_vars, c("DepDelay", "TaxiOut"))
```

## Arrival Performance
```{r}
df[, arrive_df[[1]]] %>% head(2)

# will not include ArrDelayMinutes as equivalent to ArrDelay: with negatives instead of zeros 
nominal_vars = c(nominal_vars, c("ArrDel15", 
                                 "ArrivalDelayGroups", "ArrTimeBlk"))
time_vars = c(time_vars, c("WheelsOn", "CRSArrTime", "ArrTime"))
num_vars = c(num_vars, c("TaxiIn", "ArrDelay"))
```

## Cancellations and Diversions
```{r}
df[, canc_div_df[[1]]] %>% filter(Cancelled == 1)

# Update cancellation code with factor labels from documentation
df$CancellationCode = ifelse(df$CancellationCode == "A", "CarrierDelay",
       ifelse(df$CancellationCode == "B", "WeatherDelay", 
              ifelse(df$CancellationCode == "C", "NASDelay", 
                     ifelse(df$CancellationCode == "D", "SecurityDelay", NA))))

# recall Duplicate feature only has one response, do not include column
nominal_vars = c(nominal_vars, 
                 c("Cancelled", "CancellationCode", "Diverted"))
```

## Flight Summaries
```{r}
df[, summary_df[[1]]] %>% head(2)

# recall Flights feature only has one response, do not include column
nominal_vars = c(nominal_vars, c("DistanceGroup"))
num_vars = c(num_vars, c("CRSElapsedTime", "ActualElapsedTime", "AirTime", "Distance"))
```

## Cause of Delay
```{r}
df[, delay_cause_df[[1]]] %>% na.omit() %>% head(2)

num_vars = c(num_vars, c("CarrierDelay","WeatherDelay","NASDelay",
                         "SecurityDelay","LateAircraftDelay"))
```

## Gate Return Information at Origin Airport 
```{r}
df[, gate_return_df[[1]]] %>% na.omit() %>% head(2)


colSums(is.na(df[, gate_return_df[[1]]])) / nrow(df[, gate_return_df[[1]]])
# Gate return features are missing majority of values, do not include columns
```

## Diverted Airport Information
```{r}
# diverted_air df
df[, diverted_air_df[[1]]] %>% head(2)

(colSums(is.na(df[, diverted_air_df[[1]]])) / nrow(df[, diverted_air_df[[1]]]))*100
# most Diverted features are missing majority of values, do not include columns
```

# Fix column types 
```{r}
# create function to turn integer column into time with hour:min format

fix_time_format = function(time_data) {
  # make sure to fix time in FlightDate before running
  helper_df = df %>% select(FlightDate, all_of(time_data))
  pad_time = str_pad(df[, time_data], 4, pad = "0")
  # add hour and minute columns
  helper_df$Hour = as.numeric(substr(pad_time,start=1,stop=2))
  helper_df$Minute = as.numeric(substr(pad_time,start=3,stop=4))

  # combine date and time columns
  formatted_time_df <- helper_df %>%
    mutate(datetime = FlightDate %>%
             add_hours(Hour) %>%
             add_minutes(Minute), .keep = "unused")
  
  return(formatted_time_df$datetime)
}
```

```{r}
# correct factor variable types
df[, "FlightDate"] = as.POSIXct(df[, "FlightDate"], format = "%m/%d/%Y %H:%M") #only date col
df[, time_vars] = lapply(time_vars, function(x) fix_time_format(x))
df[, nominal_vars] = lapply(nominal_vars, function(x) as.factor(df[[x]]))
df[, num_vars] = lapply(num_vars, function(x) as.numeric(df[[x]]))
```

# Construct complete data frame 
```{r}
clean_df = df %>% 
  # select variables of interest in the same order
  select(all_of(colnames(df)[colnames(df) %in% 
                               c(time_vars, date_vars, num_vars, nominal_vars)]))
```

# save clean data
```{r}
saveRDS(clean_df, file = "Data/clean_df.Rds")
```
