---
title: "Analysis"
author: "Jaqueline Ortiz"
date: "2023-04-27"
output: html_document
---

# 1. Introduction
Insert here

## 1.1 Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(rvest)
library(stringr) #pad
library(clock) #under_minutes
# only look at march 2021 for now
#df_21 = read_csv("C:/Users/jacki/Downloads/Combined_Flights_2021.csv")
```

## 1.2 Load data
```{r}
df = readRDS(file = "clean_df.rds")
```

## 1.3 Inspect data 

### 1.3.1 Dataset structure
```{r}
str(df)
```

### 1.3.2 Dataset dimensions
```{r}
dim(df)
```

### 1.3.3 Column names
```{r}
colnames(df)
```

### 1.3.4 Missing values
```{r}
# explore missing values NAs
missing_df = data.frame("var_name" = colnames(df),
                    "missing_pct" = colSums(is.na(df)) / nrow(df))

# plot percentage missing for each column
missing_df %>% 
  ggplot(aes(x=reorder(var_name,-missing_pct),y=missing_pct)) +
  geom_bar(stat="identity",fill="blue")+
  coord_flip()+theme_bw()
```

# 2. Feature engineering

## 2.1 Take closer look at columns with many missing 
```{r}
missing_df %>% filter(missing_pct >= 0.5)
# missing_df %>% filter(missing_pct > 0 & missing_pct <= 0.5)
```
* We see around 90% of observations are missing the columns: CarrierDelay, WeatherDelay, NASDelay, SecurityDelay, LateAircraftDelay. We can create a nominal variable that identifies delayed flights' largest delay cause, instead of displaying minutes delayed by each cause

## 2.2 Delay cause
```{r}
# create data subset in order to combine delay cause columns into one
delayed_DF = df %>%
  select(c(CarrierDelay, WeatherDelay, NASDelay, 
           SecurityDelay, LateAircraftDelay)) 

# find number of causes for each delayed flight 
delayed_DF_num = matrix(as.numeric((delayed_DF  %>% na.omit()) >= 15), 
                       nrow = nrow(delayed_DF %>% na.omit()), ncol = ncol(delayed_DF))
delay_cause_count = rowSums(delayed_DF_num)
# print % of flights classified counts over delayed only to create delay cause pct (non-delayed contain NAs in cells)
(table(delay_cause_count) / nrow(delayed_DF_num)) * 100
```
* Around 84.49% of flights were classified as delayed for one reason, thus we can create a more meaningful variable that identifies flight's cause of largest delay.

```{r}
# identify cause of longest delay
longest_delay = pmax(delayed_DF[[1]], delayed_DF[[2]], 
                                delayed_DF[[3]], delayed_DF[[4]], delayed_DF[[5]])

# create DelayCode variable to identify flight's longest delay cause 
df$DelayCode = as.factor(ifelse(
  longest_delay == delayed_DF[[1]], "CarrierDelay",
       ifelse(longest_delay == delayed_DF[[2]], "WeatherDelay", 
              ifelse(longest_delay == delayed_DF[[3]], "NASDelay",
                ifelse(longest_delay == delayed_DF[[4]], "SecurityDelay",
                       ifelse(longest_delay == delayed_DF[[5]], "LateAircraftDelay", NA))))))

# remove each delay cause variable
#df = df %>%
#  select(-c(CarrierDelay, WeatherDelay, NASDelay, 
#           SecurityDelay, LateAircraftDelay)) 
```

## 2.3 flight delay status
```{r}
## create variable that identifies flight performance

# ArrDel15 identifies if a plane arrived on-time (0) or delayed (1). 
# Currently cancelled and diverted flights have NA, want to update values

# add new levels to column
levels(df$ArrDel15) <- c(levels(df$ArrDel15),"2", "3")
# find index of flight types and update column value
df[grep(1, df$Cancelled), "ArrDel15"] = 2
df[grep(1, df$Diverted), "ArrDel15"] = 3

# create combined status variable
df$status = factor(df$ArrDel15, levels = 0:3, 
              labels = c("ontime", "delayed", "cancelled", "diverted"))

# remove previous status columns for each each class
#df = df %>% select(-c(ArrDel15, Cancelled, Diverted))
```

# 3. EDA 

## 3.1 Response variables

The response variable, status, identifies the performance of a departed flight. 
The percentage of flights that depart on-time is 88.50, depart delayed is 10.06, get diverted is 1.26, and get cancelled is 0.17. We can do some exploration into other status performance variables: Delayed, Cancelled. 

### 3.1.1 Status
```{r}
## create plot to show flight status distribution
total_status = sapply(100*table(df$status)/nrow(df), 
                       function(x) round(x, 2))
status_tot_df = data.frame(status = names(total_status),
                          percent = total_status)
status_tot_df$row = 1 #placeholder row
# flight status summary visualization 
ggplot(status_tot_df, aes(x = row,y = percent, fill = status)) +
    geom_bar(stat="identity") + 
    geom_text(aes(label = percent), 
        position = position_stack(vjust = 0.5))
```

### 3.1.2 Delayed flights

#### 3.1.2.1 Frequency of delay time 
```{r warning=FALSE}
# plot histogram of ArrDelayGroups
par(mfrow = c(1, 2))
delayed = df %>% filter(status == "delayed")
ggplot(delayed, aes(x=ArrivalDelayGroups)) + 
  geom_histogram(stat="count") + 
  ggtitle("Frequency of delay time grouped in 15 min increments")
ggplot(delayed, aes(x=ArrivalDelayGroups)) + 
  geom_histogram(stat="count") + 
  ggtitle("Frequency of delay time grouped in 15 min increments for delayed flights")
```
